---
title: ""
author: "Srishti Mutha, Sokna Kry, Samriddha Ghosh, Nagarajan Vaidya Subramanian"
date: "Spring 2023"
output:
  word_document: default
  pdf_document: default
geometry: margin=2.54cm
fig_caption: true

---
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
# Set your working directory
getwd()
# Load your packages
install.packages("knitr")
library(knitr)
install.packages("psych")
library(psych)
install.packages("tidyverse")
library(tidyverse)

# Set your ggplot theme
my_theme <- theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.title = element_text(colour = "#440000"),
    plot.title = element_text(colour = "blue",
                              hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(colour = "#4169e1")
  )
theme_set(my_theme)
# Load your datasets
#Source of data is eGRID 2021, accessible from EPA's website at:
#https://www.epa.gov/system/files/documents/2023-01/eGRID2021_data.xlsx
egrid2021 <- readxl::read_xlsx(path = "./Data/Raw/eGRID2021_data.xlsx",
                               sheet = "ST21",
                               skip = 1)

#eGRID is a huge dataset with several data fields and we will only be using a subset of them.
#so this next line is to subset the database and select only those fields that we will be working with.
egrid_subset <- egrid2021 %>%
  select(PSTATABB, STNAMEPCAP, STNGENAN, STNOXAN, STSO2AN, STCO2AN, STCO2EQA, STGENACL, STGENAOL, STGENAGS, STGENANC, STGENAHY, STGENABM, STGENAWI, STGENASO, STGENAGT, STGENAOF, STGENATR, STGENATN)

#Creating a new blank column to represent the region that each state is from
#based on the classification system used by Bureau of Economic Analysis
egrid_subset <- egrid_subset %>%
  add_column(.before = "PSTATABB",
             Region = "blank")

#Source for the following classification of states into regions:
#https://apps.bea.gov/regional/docs/msalist.cfm?mlist=2
NewEngland <- c("CT", "ME", "MA", "NH", "RI", "VT")
MidEast <- c("DE", "DC", "MD", "NJ", "NY", "PA")
GreatLakes <- c("IL", "IN", "MI", "OH", "WI")
Plains <- c("IA", "KS", "MN", "MO", "NE", "ND", "SD")
SouthEast <- c("AL", "AR", "FL", "GA", "KY", "LA", "MS", "NC", "SC", "TN", "VA", "WV")
SouthWest <- c("AZ", "NM", "OK", "TX")
RockyMountain <- c("CO", "ID", "MT", "UT", "WY")
FarWest <- c("CA", "HI", "NV", "OR", "WA")
NonContiguous <- c("AK", "PR") #not a classification used by BEA but one that is useful for analysing this dataset

egrid_subset$Region[egrid_subset$PSTATABB %in% NewEngland] <- "New England"
egrid_subset$Region[egrid_subset$PSTATABB %in% MidEast] <- "Mid East"
egrid_subset$Region[egrid_subset$PSTATABB %in% GreatLakes] <- "Great Lakes"
egrid_subset$Region[egrid_subset$PSTATABB %in% Plains] <- "Plains"
egrid_subset$Region[egrid_subset$PSTATABB %in% SouthEast] <- "South East"
egrid_subset$Region[egrid_subset$PSTATABB %in% SouthWest] <- "South West"
egrid_subset$Region[egrid_subset$PSTATABB %in% RockyMountain] <- "Rocky Mountain"
egrid_subset$Region[egrid_subset$PSTATABB %in% FarWest] <- "Far West"
egrid_subset$Region[egrid_subset$PSTATABB %in% NonContiguous] <- "Non-Contiguous"

egrid_subset$Region == "blank" #verifying that all states have been categorised

write.csv(egrid_subset, file = "./Data/Processed/egrid2021_st21_processed.csv")

```


# Rationale and Research Questions
The Environment Protection Agency of the US Government publishes a large dataset called the Emissions and Generation Resource Integrated Database (eGRID for short). This database describes various technical and environmental parameters of all electric power plants in the US, and is published annually. This project aims to use eGRID 2021 (which was published this year in 2023) to answer the following research questions:

<<<<<<< HEAD
1. What is the distribution across the US of electricity generation from renewable and non-renewable energy sources?
2. What is the distribution of supply from the different renewable energy sources?
3. Some summary statistics of generators of all generators, in terms of fuel source, nameplate capacity, annual net generation, and pollutants emitted;
4. Emissions from different states/regions from these generators
=======
1. What is the distribution of electricity generation across the Bureau of Economic Analysis regions in the US, from both the renewable and non-renewable energy sources?
2. What is the distribution of generation from each type of renewable energy sources across Bureau of Economic Analysis regions in the US?
3. Greenhouse gas emissions across Bureau of Economic Analysis regions in the US from electricity generators. 
4. Summary statistics of generators across the Bureau of Economic Analysis regions in the US, in terms of fuel source, nameplate capacity, annual net generation, and pollutants emitted?
5. Is there correlation between energy generators and pollutants emitted? 


These analyses will be performed by aggregating the states into categories as done by the Bureau of Economic Analysis which divides the states into 8 regions. We have added an additional "Non-Contiguous" category to represent states in the e-GRID dataset that are not part of mainland USA.

\newpage

# Dataset Information
The Environment Protection Agency of the US Government publishes a large dataset called the Emissions and Generation Resource Integrated Database (eGRID for short). This database describes various technical and environmental parameters of all electric power plants in the US, and is published annually.

\newpage

# Exploratory Analysis 

```{r summary.stats}

#summary stats of total nameplate capacity and annual generation per region
summary.plants <- egrid_subset %>%
  select(Region, STNAMEPCAP, STNGENAN) %>%
  rename("State nameplate capacity (MW)" = STNAMEPCAP,
         "State annual net generation (MWh)" = STNGENAN) %>%
  pivot_longer(cols = c("State nameplate capacity (MW)",
                        "State annual net generation (MWh)"),
               names_to = "variable", values_to = "value") %>%
  group_by(Region, variable) %>%
  summarise(obs = length(value[!is.na(value)]),
            min = round(min(value, na.rm = TRUE), 3),
            max = round(max(value, na.rm = TRUE), 3),
            mean = round(mean(value, na.rm=TRUE), 3),
            median = round(median(value, na.rm=TRUE), 3),
            sd = round(sd(value, na.rm=TRUE), 3),
            skewness = round(skew(value,na.rm = TRUE), 3))
summary.plants

#summary stats of region-wise emission of different pollutants and the CO2 equivalents
STNOXAN, STSO2AN, STCO2AN, STCO2EQA, 


#summary stats of region-wise breakdown of generation by fuel type
STGENACL, STGENAOL, STGENAGS, STGENANC, STGENAHY, STGENABM, STGENAWI, STGENASO, STGENAGT, STGENAOF

```


\newpage

# Analysis



## Question 1: What is the distribution of electricity generation across the Bureau of Economic Analysis regions in the US, from both the renewable and non-renewable energy sources?
```{r}

```

## Question 2: What is the distribution of generation from each type of renewable energy sources across Bureau of Economic Analysis regions in the US?
```{r}

```

## Question 2: 

```{r}


```

## Question 3: 
```{r}

```



## Question 4: Summary statistics of generators across the Bureau of Economic Analysis regions in the US, in terms of fuel source, nameplate capacity, annual net generation, and pollutants emitted?
```{r}
# Boxplot
ggplot(egrid_subset, aes(x = Region, y = STCO2AN)) + 
  geom_boxplot()

ggplot(egrid_subset, aes(x = Region, y = STSO2AN)) + 
  geom_boxplot()

ggplot(egrid_subset, aes(x = Region, y = STCO2EQA)) + 
  geom_boxplot()

ggplot(egrid_subset, aes(x = Region, y = STNOXAN)) + 
  geom_boxplot()


ggplot(egrid_subset, aes(x = Region, y = STGENASO)) + 
  geom_boxplot()


```

## Question 5: Is there correlation between energy generators and pollutants emitted? 
Ho: There is no correlation between energy generators and pollutants emitted
H1: There is correlation between energy generators and pollutants emitted

```{r}
#T.Test

T_GENOX <- lm(egrid_subset$STNGENAN ~ egrid_subset$STNOXAN)
summary(T_GENOX)

T_GESO2 <-lm(egrid_subset$STNGENAN ~ egrid_subset$STSO2AN)
summary(T_GESO2)

T_GECO2 <- lm(egrid_subset$STNGENAN ~ egrid_subset$STCO2AN)
summary(T_GECO2)

T_GECO2EQA <- lm(egrid_subset$STNGENAN ~ egrid_subset$STCO2EQA)
summary(T_GECO2EQA)
```


\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
